\newpage
\section{Service Architecture Model}
\label{sec:samModel}
This section define precisely all the Service Architecture Model concepts and assumptions.

To remark the use of SAM concepts on the text Start Case is used for concept names.

Defined sets and relations are related to one Architecture defined using Service Architecture Model.

\subsection{Types and Interface}

A object oriented approach is taken to define Types and Interfaces.

\begin{defi}[Basic Type]
  A predefined set of primitive types.
\end{defi}

A concrete list of accessible Basic Types is related to a given implementation of SAM. Bindings to specific programming languages may define mappings between provided Basic Types on SAM implementation and language types.

\begin{defi}[Type]
  Basic Type and Interface Types uniquely defined by a Interface.
\end{defi}

\begin{defi}[Interface]
  Interface is a set of methods. Method signature is defined by a
\begin{itemize}
 \item method Name
 \item return Type
 \item list of arguments (list of Types)
\end{itemize}
\end{defi}

An Interface should be uniquely identified by a name.

% \begin{zalozenie}
%  \item Architecture Interfaces don't have class members.
% \end{zalozenie}


The set of Interfaces is denoted by $\mathcal{I}$. The set of Types is denoted by $\mathcal{T}$. The set of Basic Types is denoted by $\mathcal{B}\subset\mathcal{T}$.

The mapping between Interface and corresponding Interface Type is denoted by function $t$
\begin{eqnarray}
t:\mathcal{I}\rightarrow\mathcal{T}
\end{eqnarray}
and implies also a function $\tilde{t}$ between Interfaces sets to Types sets defined as:
\begin{eqnarray}
\tilde{t}:2^\mathcal{I}\rightarrow 2^\mathcal{T} \quad
 \tilde{t}(\{i_1,\dots,i_\alpha\}) := \{t(i_1),\dots,t(i_\alpha) \}
\end{eqnarray}

\begin{defi}[Signature Relation]
  Interface `$i$' is signature related to Type `$t$' if a method signature of `$i$' contains `$t$'. This relation is denoted by set `$\mathring{s}$':
\begin{eqnarray}
\mathring{s}\subset \mathcal{I}\times \mathcal{T}
\end{eqnarray}
\end{defi}

\begin{defi}[Signature Maps]
The Signature Map is the function $s$
\begin{eqnarray}
&s:\mathcal{I}\rightarrow 2^\mathcal{T}&
\end{eqnarray}
given by the statement:
\begin{eqnarray}
&(i,t) \in \mathring{s} \Rightarrow t\in s(i) &\nonumber
\end{eqnarray}

A set version of the Signature Map is defined by function $\tilde{s}$
\begin{eqnarray}
&\tilde{s}:2^\mathcal{I}\rightarrow 2^\mathcal{T}& 
\end{eqnarray}
defined as 
\begin{eqnarray}
&\tilde{s}(\{i_1,\dots,i_\alpha\}) := s(i_1) \cup \dots \cup s(i_\alpha) &
\end{eqnarray}

\end{defi}

\begin{zalozenie}
 \item For a given Architecture, Signature methods of Interfaces use only Basic Types or Types defined by Interfaces belonging to the given Architecture.
\end{zalozenie}

\subsection{Domain Specification Model}

\begin{defi}[Domain Specification]
  A named set of Interfaces.
\end{defi}

\begin{zalozenie}
 \item Domain Specifications on Architecture are mutually disjoint.
\end{zalozenie}

% The set of Interfaces is denoted by $\mathcal{I}$. The set of Types related to $\mathcal{I}$ is denoted by $\mathcal{T}$. The set of Basic Types is denoted by $\mathcal{B}\subset\mathcal{T}$.

The set of Domain Specifications is denoted by $\mathcal{D}$. Because for $d \in \mathcal{D}$ we have $d \subset 2^\mathcal{I}$, function $\tilde{t}$ and $\tilde{s}$ are well defined on $\mathcal{D}$.

\begin{defi}[Domain Dependency Relation]
  Domain Specification $d_1$ \textbf{depends on} Domain Specification $d_2$ $[(d_1,d_2)\in \delta]$, if a Type related to a Interface belonging to $d_2$ is used on a method signature of a Interface belonging to $d_1$
\begin{eqnarray}
 &\delta \subset \mathcal{D} \times \mathcal{D} &\\
&\tilde{t}(d_2) \cap \tilde{s}(d_1) \neq \emptyset  \Rightarrow (d_1,d_2) \in \delta  &\nonumber
\end{eqnarray}
\end{defi}

\begin{zalozenie}
 \item Domain Dependency Relation $\delta$ is acyclic.
\end{zalozenie}

Simple Domain Specification is a Domain Interfaces that use in methods signature only Basic Types and own Types, precisely:
\begin{defi}[Simple Domain Specification]
  Domain Specification $d$ such that
\begin{eqnarray}
\tilde{s}(d) \subset \mathcal{B} \cup \tilde{t}(d)
\end{eqnarray}
\end{defi}

Definition implies that Simple Domain Specification don't depend on any Domain Specification.

\begin{defi}[Model Domain Specification]
  Domain Specification that depends on other Domain Specification.
\end{defi}

\subsection{Service Specification}

\begin{defi}[Service Specification]
  A Model Domain Specification.
\end{defi}

\begin{zalozenie}
 \item Services are not used as Domain Dependencies: if $s$ is a Service Specification, then
\begin{eqnarray}
 \forall d \in \mathcal{D} \quad  (d,s) \notin \delta
\end{eqnarray}
\end{zalozenie}

The set of Service Specification is denoted by $\mathcal{S}$, we have $\mathcal{S} \subset \mathcal{D}$.

\subsection{Taxonomy and Category}

\begin{defi}[Taxonomy]
  Equivalence relation of Services Specifications
\begin{eqnarray}
 \tau \subset \mathcal{S}  \times \mathcal{S} 
\end{eqnarray}
\end{defi}

\begin{defi}[Category]
  Equivalence class of Taxonomy.
\end{defi}

The set of Categories is denoted by $\Gamma$. Taxonomy projection is denoted by
\begin{eqnarray}
cat: \mathcal{S} \rightarrow \Gamma
\end{eqnarray}

\subsection{Service Implementation}

Service Implementation definition is based on Dependency Injection Model presented on section \ref{sec:ioc}.

\begin{defi}[Architecture Annotations]
  A set of Binding Annotations.
\end{defi}

\begin{defi}[Service Implementation]
  A Binding Provider on Architecture Annotations for all the Types related to a Service Specification.
\end{defi}

The set of Service Implementations is denoted by $\mathcal{P}$. Function $spec$
\begin{eqnarray}
 spec: \mathcal{P}  \rightarrow \mathcal{S}
\end{eqnarray}
assigns Service Implementation to realized Services Specifications.


\begin{defi}[Injection Relation]
  Service Implementation $p$ \textbf{injects} Service Specification $s$ $[(p,s)\in\kappa]$ if the following chain can be build:
\begin{eqnarray}
\mathcal{P} \ni p \xrightarrow[\text{contain Class}]{} c 
\xrightarrow[\text{with Injection Point}]{} ip
\xrightarrow[\text{matching Key}]{} k  
\xrightarrow[\text{of Type}]{} a \nonumber \\ 
a \xrightarrow[\text{related to Interface}]{t(i) = a} i
\xrightarrow[\text{belonging to Service Specification}]{i\in s} s \in \mathcal{S} \nonumber
\end{eqnarray}
\begin{eqnarray}
 \kappa \subset \mathcal{P} \times \mathcal{S}
\end{eqnarray}
\end{defi}

Initial four relations on definition of Injection Relation are defined in Dependency Injection Model section \ref{sec:ioc}. The last two relations are defined by logical condition over the arrows.

\begin{defi}[Set of Injected Service Specifications]
  of Service Implementation $p$ is 
\begin{eqnarray}
&injected : \mathcal{P} \rightarrow 2^\mathcal{S}& \\
&injected(p) := \{ s : (p,s)\in \kappa \} \subset \mathcal{S} &\nonumber
\end{eqnarray}
\end{defi}

\subsection{Category Accessibility Relation}

\begin{defi}[Category Accessibility Relation]
  A predefined relation
\begin{eqnarray}
\pi \subset \Gamma \times \Gamma
\end{eqnarray}
\end{defi}

Category $c_2$ \textbf{is accessible} to Category $c_1$ if $(c_1,c_2) \in \pi$.

As determined by the next assumption, Category Accessibility Relation $\pi$ restrict injection of Services on Implementations. 

\begin{zalozenie}
 \item $ \forall s_x \in injected(p) \quad  (cat(spec(p)),cat(s_x)) \in \pi$.
\end{zalozenie}

This mean that if the following conditions are satisfied
\begin{itemize}
\item Implementation $p$ implements Service $s$ [$spec(p) = s $]
\item Implementation $p$ injects Service $s_x$ [$ s_x \in injected(p) \equiv (p,s_x)\in\kappa$ ]
\item Service $s$ belongs to Category $C$ [$s\in C$]
\item Service $s_x$ belongs to Category  $C_x$ [$s_x\in C_x$]
\end{itemize}
 then it is required that Category $C_x$ is accessible to Category $C$ [$(C,C_x)\in \pi $].

\subsection{Prototype and Test Specifications}

For development of Service Implementations it is necessary to have access to external services for local execution and testing.

\begin{defi}[Prototype Implementation]
  is a Service Implementation that don't inject any Service Specification.
\end{defi}

Prototypes can be executed without any other resource, as they don't inject any service.

\begin{defi}[Test Specification]
  is a Binding Provider that injects only one Service Specification.
\end{defi}

Test Specifications can inject a Service Implementation and execute test using the Service Specification Interfaces. Test Specifications may implement a Service Specification to control the execution of tests and provide results in a specific format, but this is a per Architecture choice.

Test Specifications may be treated as formal specifications for a Service Specification.

